---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r data,echo=TRUE}
if (!file.exists('activity.csv')) {
  unzip(zipfile = "activity.zip")
}

activityData <- read.csv(file="activity.csv", header=TRUE)
```


## What is mean total number of steps taken per day?
```{r mean_steps,echo=TRUE}
totalSteps <- aggregate(steps ~ date, activityData, FUN=sum)

hist(totalSteps$steps,
     main = "Total Steps per Day",
     xlab = "Number of Steps")

meanSteps <- mean(totalSteps$steps, na.rm = TRUE)
medSteps <- median(totalSteps$steps, na.rm = TRUE)
```
Mean Number of Steps Taken per Day `r meanSteps`,
Median Number of Steps Taken per Day `r medSteps`

## What is the average daily activity pattern?
```{r echo=TRUE}
library(ggplot2)
meanStepsByInt <- aggregate(steps ~ interval, activityData, mean)
ggplot(data = meanStepsByInt, aes(x = interval, y = steps)) +
  geom_line() +
  ggtitle("Average Daily Activity Pattern") +
  xlab("5-minute Interval") +
  ylab("Average Number of Steps") +
  theme(plot.title = element_text(hjust = 0.5))

maxInt <- meanStepsByInt[which.max(meanStepsByInt$steps),]
```



## Imputing missing values
```{r echo=TRUE}
missingVals <- is.na(activityData$steps)
missCount <- sum(missingVals)
```

There are `r missCount` missing values. I will replace these missing values with the 5-day average of that respective interval.

```{r echo=TRUE}
imp_activityData <- transform(activityData,
                              steps = ifelse(is.na(activityData$steps),
                                             meanStepsByInt$steps[match(activityData$interval, 
                                                                        meanStepsByInt$interval)],activityData$steps))

impStepsByInt <- aggregate(steps ~ date, imp_activityData, FUN=sum)
hist(impStepsByInt$steps,
     main = "Imputed Number of Steps Per Day",
     xlab = "Number of Steps")
     
impMeanSteps <- mean(impStepsByInt$steps, na.rm = TRUE)
impMedSteps <- median(impStepsByInt$steps, na.rm = TRUE)
diffMean = impMeanSteps - meanSteps
diffMed = impMedSteps - medSteps
diffTotal = sum(impStepsByInt$steps) - sum(totalSteps$steps)
```

There is a difference of `r diffMean` in the mean steps of the two dataset. There is a difference of `r diffMed` in the median steps of the two dataset. There is a difference of `r diffTotal` in the total steps of the two dataset.

## Are there differences in activity patterns between weekdays and weekends?
```{r echo=TRUE}
DayType <- function(date) {
  day <- weekdays(date)
  if (day %in% c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday'))
      return ("weekeday")
  else if (day %in% c('Saturday', 'Sunday'))
      return ("weekend")
  else
      stop ("Invalid Date Format.")
}
imp_activityData$date <- as.Date(imp_activityData$date)
imp_activityData$day <- sapply(imp_activityData$date, FUN = DayType)

meanStepsByDay <- aggregate(steps ~ interval + day, imp_activityData, mean)
ggplot(data = meanStepsByDay, aes(x = interval, y = steps)) + 
  geom_line() +
  facet_grid(day ~ .) +
  ggtitle("Average Daily Activity Pattern") +
  xlab("5-minute Interval") +
  ylab("Average Number of Steps") +
  theme(plot.title = element_text(hjust = 0.5))
```
